State of the Practice for Lattice Boltzmann Method Software

### 2022-03-02 StateOfPractice/Papers/LatticeBoltz_SOP-arXiv/LatticeBoltz_SOP.tex
Edits up to and including Section 10.6 Augment Theory Manual to Include Requirements Information

\begin{itemize}
	\item Explicitly state appropriate fluid dynamics problems that the software
	is designed to model and explicitly state the limits of the software. This
	is done in the ESPResSo user guide.
	\item Provide documentation that details the background theory, or provide a
	reference to such information. This was done by several top ranked packages,
	including ESPResSo, Ludwig, LUMA, and OpenLB.
	\item Identify expected user characteristics. LIMBES, Ludwig, laboetie, and
	Palabos all did this. Explicit user characteristics are important so the
	application and documentation can be pitched toward the right user
	background \citep{smith2007requirements}.
	\item Keep all documentation in one location. This was done by all top
	ranked packages.
	\item Use documentation generators, like Doxygen, as done (for instance) by
	OpenLB.  More details on the packages that use generators is given in
	Sections~\ref{SecSurfCorrectAndVerifiab} and~\ref{Sec_CompareTools}.
	\item Provide an explanation of theoretical principles through the user interface.
	\item Add guards in the code to test user inputs for compatibility before
	proceeding with the calculations.
	\item Include documentation on frequently asked questions (Section~\ref{Sec_CompareArtifacts}).
\end{itemize}

@@

Addresses lack of development time (\ppref{P_LackDevTime}) by increasing the
number of developers.

Maintainability. Have a defined process for accepting contributions, and make
public documentation for making contributions to the project. The top five
ranked software packages include information on how to contribute.

Papers that talk about how to do open source. \citep{Jalan2016}

Get more developers by better documentation, tutorials, etc.

Combine existing teams?

@@

- say what PPs are addressed
- further steps to address the oracle problem - borrow from SE for HEP paper.
- anything in Smith 2016?
- introduce list on page 11 of HEP paper, add repeatability point

In the absence of a test oracle, the following
techniques can be used to build system tests \citep{Smith2016}:

@@

\item Test cases can be selected that are a subset of the real problem for which
  a closed-form solution exists.  When using this approach, confidence in the
  actual production code can only be built if it is the same code used for
  testing; that is, nothing is gained if a separate, simpler, program is written
  for testing the special cases.
\item Verification test cases can be created by assuming a solution and using
  this to calculate the inputs.  For instance, for a linear solver, if $A$ and
  $x$ are assumed, $b$ can be calculated as $b = A x$.  Following this, $A x^* =
  b$ can be solved and then $x$ and $x^*$, which should theoretically be equal,
  can be compared.  In the case of solving Partial Differential Equations
  (PDEs), this approach is called the Method of Manufactured Solutions
  \citep{Roache1998}.
\item Most scientific software uses floating point arithmetic, but for testing
  purposes, the slower, but guaranteed correct, interval arithmetic
  \citep{Hickey2001} can be employed. The faster floating point algorithm can
  then be verified by ensuring that the calculated answers lie within the
  guaranteed bounds.
\item Verification tests should include plans for convergence studies.  The
  discretization used in the numerical algorithm should be decreased (usually
  halved) and the change in the solution assessed.
\item Confidence can be built in a numerical algorithm by comparing the results
  to another program that overlaps in functionality.  If the test results do not
  agree, then one, possibly both of the programs is incorrect.
\item The verification plan should also include test plans for nonfunctional
  requirements, like accuracy, performance and portability, if these are
  important implementation goals.  Performance tests can be planned to describe
  how the software responds to changing inputs, such as problem size, condition
  number etc.  Verification plans can include relative comparisons between
  the new implementation and competing products \citep{Smith2006}.
\item In cases where a test oracle is not available, as is usually the case for
  scientific software, \emph{metamorphic} testing can be used.  ``This technique
  operates by checking whether the program behaves according to a certain set of
  properties called metamorphic relations.  A metamorphic relation is a
  relationship between multiple inputs and output pairs of the program.''
  \citep{KanewalaAndLundgren2016}.
\end{itemize}

@@

Point to Section~\ref{Sec_CompareArtifacts}. Requirements documentation is
feasible, as shown by the prevalence of theory documents in LBM software.
Table~\ref{artifactspresent} shows that theory documentation is common.  We
observed 17 of 24 packages had at least some theory documentation. Templates for
scientific requirements, like \citet{SmithEtAl2007, SmithAndLai2005}, show
theory is a significant part of the requirements documentation. With the
addition of some extra information, the theory documents can be transformed into
requirements specification.  The key extra information includes explicitly
stating user characteristics, explicitly stating how the user interacts with the
software in terms of input data requirements, and listing likely and unlikely
changes.  Explicit statements about likely changes are invaluable in the design
stage, since they provide developers guidance on how general the software needs
to be.  For those wishing to maximize the value of a requirement specification,
information could be added like prioritizing the nonfunctional requirements (to
show the relative importance between qualities like portability, reliability and
performance) and traceability information (to show the consequences of any
changes to the assumptions).

- say what PPs are addressed
- previous section discusses challenges with requirements and not popular (maybe
move that section here) - say what information to incorporate, low hanging fruit
- return on investment
- maybe borrow from SE for HEP
- toward goal of replicability \citep{BenureauAndRougier2017}.

Addresses lack of development time (\ppref{P_LackDevTime}) by getting developers
up to speed more quickly.  Reducing the amount of time needed to train new
developers.

Surface Correctness and Verifiability. Use a requirements
specification document. This is suggested in SCS literature, and several of
the top ranked packages had such a document or reference to theory manuals.
A potential template is presented in \citep{smith2005new}.

@@

>>>>

\begin{itemize}
	\item Explicitly state appropriate fluid dynamics problems that the software
	is designed to model and explicitly state the limits of the software. This
	is done in the ESPResSo user guide.
	\item Provide documentation that details the background theory, or provide a
	reference to such information. This was done by several top ranked packages,
	including ESPResSo, Ludwig, LUMA, and OpenLB.
	\item Identify expected user characteristics. LIMBES, Ludwig, laboetie, and
	Palabos all did this. Explicit user characteristics are important so the
	application and documentation can be pitched toward the right user
	background \citep{SmithEtAl2007}.
	\item Keep all documentation in one location. This was done by all top
	ranked packages.
	\item Use documentation generators, like Doxygen, as done (for instance) by
	OpenLB.  More details on the packages that use generators is given in
	Sections~\ref{SecSurfCorrectAndVerifiab} and~\ref{Sec_CompareTools}.
	\item Provide an explanation of theoretical principles through the user interface.
	\item Add guards in the code to test user inputs for compatibility before
	proceeding with the calculations.
	\item Include documentation on frequently asked questions (Section~\ref{Sec_CompareArtifacts}).
\end{itemize}

@@

In our interviews with developers we heard they would like to have additional
software contributors.  More developers would help with addressing the lack of
development time pain point (\ppref{P_LackDevTime}).  In investigating advice on
increasing the number of contributors, we found that the advice usually starts
with following best practices, including providing artifacts like those
discussed in Section~\ref{Sec_CompareArtifacts} such as a contributor
guidelines, a clear code of conduct and high quality code and documentation.
This advice is logical, but in our assessment of LBM software we found examples
of projects that already follow many best practices, but still have small
development teams. Attracting developers apparently requires more than just
building a good product.

Some potential additional ideas for growing the number of contributors are
as follows:

\begin{itemize}
	\item Clearly identify issues that are an appropriate starting points for
	new developers \citep{Garcia2016, Jalan2016, Proffitt2017}.
	\item Given that in most projects developers were once users
	\citep{McQuaid2018}, recruit future developers from the current set of
	users.
	\item Create templates for pull requests and issues \citep{Jalan2016}.
    \item Welcome all kinds of contributions, not just code.  Non-code
    contributions include documentation, fixing typos, issue reporting and test
    cases \citep{Jalan2016, Proffitt2017}.
    \item Reward and recognize new contributors, via small rewards like stickers
    or shirts, or even just a simple shoutout or mention in a blogpost or on
    social media \citep{Jalan2016, Proffitt2017}.
    \item Recognize that open source is more about people than it will ever be
    about code \citep{Jalan2016}.
    \item Look beyond just recruiting online and seek new developers at
    conferences or other user meet ups \citep{Garcia2016}.
    \item Follow the advice of \citet{Kuchner2012} to adapt ideas from marketing
    to promote science.  Specific ideas that could be applied to open source
    projects include the following: 
    \begin{itemize}
		\item Storytelling to motivate interest in the project, where the story
		is a sequence of events and pauses for reflection \citep[p.\
		21--22]{Kuchner2012}.
		\item Invest effort in building relationships with users and with potential future developers.
		\item Brand your project with what makes it unique and special.
	\end{itemize}
	\item Consider combining existing projects to pool collective resources, and
	reduce competition between different open source projects.
\end{itemize}

@@

To address the pain point of ensuring correctness (\ref{P_Correctness}),
Section~\ref{SecReproducibility} outlined techniques that LBM developers use to
circumvent the oracle problem for testing.  Additional techniques to build
system tests are as follows \citep{Smith2016}:

@@

\item Create test cases by assuming a solution and using this solution to
  calculate the necessary inputs.  For instance, for a linear solver, if $A$ and
  $x$ are assumed, $b$ can be calculated as $b = A x$.  Following this, $A x^* =
  b$ can be solved and then $x$ and $x^*$, which should theoretically be equal,
  can be compared.  In the case of solving Partial Differential Equations
  (PDEs), this approach is called the Method of Manufactured Solutions
  \citep{Roache1998}.
\item Use interval arithmetic to build test cases.  For testing purposes, the
  slow, but guaranteed correct, interval arithmetic \citep{Hickey2001} can be
  employed. The faster floating point algorithm can then be verified by ensuring
  that the calculated answers lie within the guaranteed bounds.
\item Include convergence studies.  The discretization used in the numerical
  algorithm should be decreased (usually halved) and the change in the solution
  assessed.  This is very likely already done by LBM developers, but it did not
  come up during our interviews.
\item Use \emph{metamorphic} testing, where the program is checked to see if it
  satisfies a set of metamorphic relations, where a metamorphic relation is a
  relationship between multiple inputs and output pairs
  \citep{KanewalaAndLundgren2016}.  An example could be the relationship that
  the output flow rate increases as the input flow rate increases.
\end{itemize}

@@

As discussed in Section~\ref{Sec_CompareArtifacts}, requirements specifications
are rarely written for research software. Although a full requirements
specification has advantages \citep{SmithEtAl2007, SmithAndLai2005},
given the lack of development time (\ref{P_LackDevTime}), we recommend the
intermediate step of augmenting the existing theory manuals with some
requirements information.  Table~\ref{artifactspresent} shows that theory
documentation is common.  We observed 17 of 24 packages had at least some theory
documentation. Templates for scientific requirements, like \citet{SmithEtAl2007,
SmithAndLai2005}, show theory is a significant part of the requirements
documentation. With the addition of some extra information, the theory documents
can be transformed into requirements specification.  The key extra information
includes explicitly stating user characteristics, explicitly stating how the
user interacts with the software in terms of input data requirements, and
listing likely and unlikely changes.  Explicit statements about likely changes
are invaluable in the design stage, since they provide developers guidance on
how general the software needs to be.  For those wishing to maximize the value
of a requirement specification, information could be added like prioritizing the
nonfunctional requirements (to show the relative importance between qualities
like portability, reliability and performance) and traceability information (to
show the consequences of changes to the assumptions). Spending time to
incorporate additional information in the theory manuals should help address the
quality of existing documentation (\ref{P_Documentation})  and improve the
software's usability (\ref{P_Usability}).

@@

- requirements are needed for the goal of replicability
\citep{BenureauAndRougier2017}.
